\clearpage
\addcontentsline{toc}{chapter}{Acknowledgements}
\begin{center}
\textbf{\Large Acknowledgements}\\
\vspace{\baselineskip}
\end{center}

%Insert your acknowledgement text here
I would like to express my sincere gratitude to Dr. Kostas Papakonstantinou, my 
academic advisor and chair of my doctoral committee, for his guidance, 
encouragement and valuable assistance throughout all these years. Without 
his guidance, this work would not have come into fruition and for that I am 
deeply grateful.

Special thanks are owed to the rest members of the dissertation committee, Dr. 
Gordon Warn, Dr. Pinlei Chen, Dr. Tong Qiu, Dr. Reuben Kraft, and the former 
member Dr. Michael Hillman, for their 
participation, advice and suggestions in completing this work. I would also 
like to thank my friend and colleague Dr. Charalampos Andriotis for all his 
helpful suggestions and important contributions early on in this endeavor.

I cannot conclude this acknowledgement without expressing my deep gratitude to 
my parents, Claire and Michael, to whom this dissertation is dedicated. They  
have always been by my side with their encouragement, love and unwavering 
support, without which none of this would have been possible.

The material presented herein is based upon work supported by the U.S. National 
Science Foundation under Award No. 1634575. The support is greatly 
acknowledged. Any opinions, findings, and conclusions or recommendations 
expressed in this publication are those of the author and do not necessarily 
reflect the views of the U.S. National Science Foundation.
\clearpage
%\pagenumbering{gobble}  %remove page number on summary page
